Response time of a plasmonic distributed feedback laser in large-signal modulation regime

### 2019-03-31 DFB_Pulse_LaTeX/DFB_pulse_inProcess.tex
Modified by AZ

	The time response to an external signal is the main characteristic of optoelectronic devices which determines their maximum modulation frequency.
	The use of plasmonic structures gives the opportunity to significantly reduce the response time.
	In this paper, we study the temporal dynamics of a plasmonic distributed feedback laser consisting of a two-dimensional metallic structure coated with a layer of the active medium.
	Combining numerical simulations with an analytical evaluation we show that the response time in the large-signal modulation regime strongly depends on the size of the pump beam.
	We demonstrate that the interaction of a large number of modes of the plasmonic structure with the pumped active medium leads to the formation of a single collective mode.
	If the size of the pump beam is larger than the propagation length of electromagnetic waves in the plasmonic structure, the collective mode is localized in the pumped area.
	In this case, the response time slightly increases with the size of the pump beam.
	In the opposite case, the mode is not localized inside the pumped area and the response time grows with the narrowing of the beam.
	We demonstrate that there is an optimal size of the pump beam for which the modulation speed is a maximum, and for typical experimental parameters can achieve $1$ THz.
	The obtained result opens the way to increasing both the modulation speed and the energy efficiency of plasmonic laser devices.

@@

In our paper, we study the dynamics in the large-signal modulation regime of a plasmonic distributed feedback (DFB) laser that consists of a gold film perforated by a periodic array of holes, and is coated with a layer of an active medium.
We examine the behavior of the laser for different sizes of the pump beam.
Combining numerical simulations of the laser dynamics within the framework developed in Ref. \cite{Zyablovsky2017approach,nefedkin2018acsphot} with an analytical evaluation, we show that the response time strongly depends on the size of the pump beam.
This dependence is due to the complex structure of the modes of the DFB laser.
The interaction of the laser modes with the pumped active medium causes the emergence of a single collective mode whose properties strongly depend on the pump.
If the pump beam is larger than the decay length of the EM field in the plasmonic structure, the collective mode is localized inside the pumped area.
The response time in this case reaches picoseconds and slightly increases with the size of the pump beam.
In the opposite case, the mode is not localized inside the pumped area and the response time grows with the narrowing of the beam.
We demonstrate that there is an optimal size of the pump beam, for which the response time reaches a minimum.
For typical experimental parameters, the optimal size of the pump beam is $\sim 15 \mu$m and the minimal response time is $\sim 1$ $\text{ps}$.
The obtained value of the response time is in agreement with a recent experiment \cite{daskalakis2018nanolett} where it has been shown that plasmonic lasers have a modulation frequency exceeding several hundred GHz in the large-signal modulation regime.
Thus, the optimization of the pump beam size makes it possible to increase the modulation speed as well as the energy efficiency of plasmonic DFB laser devices.

@@

\section*{Temporal dynamics of a two-dimensional plasmonic DFB laser}
A conventional plasmonic DFB laser consists of a periodic plasmonic structure, e.g., a metallic film perforated by an array of holes \cite{BeijnumPRL,DorofeenkoOptExp,MengLPR,nefedkin2018acsphot,TennerJOpt,TennerACSPhot,Zyablovsky2017Optimum,Melentiev2017nanolaser} or two-dimensional arrays of plasmonic nanoparticles \cite{ZhouNatNano,SchokkerPRB,daskalakis2018nanolett,SchokkerACSPhot,YangNatComm,YangACSNano,hakala2017lasing,wang2017band}, and an active medium.
The eigenmodes of such a system are Bloch modes which are distributed over the whole surface of the laser.
To overcome losses in the metal, an external pumping (optical or electrical) of the active medium is used.
In the case of optical pumping, the diameter of the pump beam is usually smaller than the size of the periodic plasmonic structure \cite{TennerACSPhot,ZhouNatNano,daskalakis2018nanolett}, see Fig.~\ref{fig1}a.
In such a case, the EM field is excited inside the pumped area and is formed by a large number of Bloch modes \cite{TennerACSPhot,nefedkin2018acsphot}.
To describe the multimode dynamics of the plasmonic DFB laser we use an approach which has been proposed in Ref. \cite{Zyablovsky2017approach} for considering ultrafast phenomena in dispersive dissipative media.
This approach is based on the following equations:
\begin{equation}\label{eq1}
\frac{{d{n_{jk}}}}{{dt}} =  - \left( {{\gamma _j} + {\gamma _k}} \right){n_{jk}} + i\left( {{\omega _j} - {\omega _k}} \right){n_{jk}} + \sum\limits_{m=1}^{N_{\text{atom}}} {\left( {{\Omega _{km}}{\varphi _{jm}} + \Omega _{jm}^*\varphi _{km}^*} \right)} 
\end{equation}
\begin{equation}\label{eq2}
\frac{{d{D_m}}}{{dt}} =  - {\gamma _D}\left( {1 + {D_m}} \right)  - 2\sum\limits_{j=1}^{N_{\text{mode}}} {\left( {{\Omega _{jm}}{\varphi _{jm}} + \Omega _{jm}^*\varphi _{jm}^*} \right)} 
\end{equation}
\begin{equation}\label{eq3}
\frac{{d{\varphi _{jm}}}}{{dt}} =  -\gamma_{jm}^{\varphi} {\varphi _{jm}} + i\left( {{\omega _j^a} - {\omega _\sigma }} \right){\varphi _{jm}} + \frac{{\Omega _{jm}^*}}{2}\left( {{D_m} + 1} \right) + \sum\limits_{l=1}^{N_{\text{mode}}} {\Omega _{lm}^*{n_{jl}}{D_m}} , 
\end{equation}
Here, $n_{jj}$ is the photon number in the $j$th Bloch mode, $n_{jk}$ is an interference term responsible for the flow of photons from the $k$th to the $j$th mode, when $j \neq k$.
$D_m$ is the population inversion of the $m$th atom and $\varphi _{jm} = - i \langle \hat{a} _j^{+} \hat{\sigma} _m \rangle$ describes the energy flow between the $m$th atom and $j$th mode.
The parameters $\omega _j$, $\gamma _j$ correspond to the real and imaginary parts of the eigenfrequencies of the modes; $\omega _{\sigma}$ and $\gamma _D$ are the atom transition frequency and the relaxation rate of the population inversion of the atoms, $\gamma _{jm}^{\varphi} = \gamma _{\sigma} + \gamma _j + \gamma _D / 2$, $\gamma _{\sigma}$ is the relaxation rate of the phase of the atom polarization.
$\Omega _{jm}$ is the Rabi constant of coupling between the $j$th mode and the $m$th atom.
These equations generalize the rate equations widely used in laser theory \cite{SiegmanLasers}.
In addition to the dynamics of the photon number in each mode and the population inversion of each atom, they describe energy flows between the active medium and the atoms, and the interference between the EM fields of the different modes, $n_{jk}$.

For a quantitative description of the temporal dynamics of a plasmonic DFB laser under pulse pumping, we consider a laser based on an Au film perforated by a periodic array of nanosized holes, see Fig.~\ref{fig1}.
Such a structure has been realized in a number of experiments \cite{BeijnumPRL,TennerJOpt,TennerACSPhot} and its parameters have been well studied \cite{BeijnumPRL,TennerJOpt,TennerACSPhot,nefedkin2018acsphot}.
We consider the response of the system to ultrashort pulse pumping of the active medium.
Assuming that the time of the external pump pulse lies in the femtosecond range and is much shorter than all characteristic times of the plasmonic DFB laser, in our model the ultrashort pulse pumping is described as the initial condition for the $D_m$ in Eqs.~(\ref{eq1})--(\ref{eq3}).

@@

\begin{figure}[h]
	\centering
	\begin{subfigure}[h]{0.32\linewidth}
		\includegraphics[width=\linewidth]{Fig3a.png}
%		\caption{}
	\end{subfigure}
	\begin{subfigure}[h]{0.32\linewidth}
		\includegraphics[width=\linewidth]{Fig3b.png}
%		\caption{}
	\end{subfigure}
	\begin{subfigure}[h]{0.32\linewidth}
		\includegraphics[width=\linewidth]{Fig3c.png}
%		\caption{}
	\end{subfigure}
	\\
	\begin{subfigure}[h]{0.32\linewidth}
		\includegraphics[width=\linewidth]{Fig3d.png}
%		\caption{}
	\end{subfigure}
	\begin{subfigure}[h]{0.32\linewidth}
		\includegraphics[width=\linewidth]{Fig3e.png}
%		\caption{}
	\end{subfigure}
	\begin{subfigure}[h]{0.32\linewidth}
		\includegraphics[width=\linewidth]{Fig3f.png}
%		\caption{}
	\end{subfigure}
	\caption{Field distribution of the collective mode when area of the pump beam $d_{\text{beam}}<d_{\text{beam}}^*$ (a) -(c) and $d_{\text{beam}}>d_{\text{beam}}^*$ (d)-(f). The time moments in (a)-(c) and (d)-(f) correspond to circles on the red curve and the black curve in Fig.\ref{fig2}a, respectively.}
	\label{fig3}	
\end{figure}

@@

The existence of an optimal size of the pump beam can be explained with the example of a single-mode model.
To justify the single-mode approximation, we calculate the distribution of the total electric field over the area of the DFB laser at different moments of time, see Fig.~\ref{fig3}.
When the diameter of the pump beam is larger than the critical value, $d_{\text{beam}} > d_{\text{beam}}^*$, the electric field is located inside the pumped region of the active medium, see Figs.~\ref{fig3}(d)--(f).
In the opposite case, $d_{\text{beam}} < d_{\text{beam}}^*$, the main part of the electric field is outside the pumped area, see Figs.~\ref{fig3}(a)--(c).
However, in both cases, the spatial distribution of the total electric field remains almost unchanged during its temporal evolution.
The time independence of the EM field distribution allows us to consider the plasmonic laser as a single mode laser.

Based on this observation, we assume that the part of the active medium pumped by an external beam effectively excites a single collective mode, which is formed by the Bloch modes of the plasmonic structure.
In this case, the system dynamics can be approximately described by the following system of equations \cite{SiegmanLasers}:
\begin{gather} 
\dot n =  - 2{\gamma _a}n + 2{N_{\text{at}}}{\Omega _R}\varphi \label{ZE1}
\\
\dot \varphi = - \left( {{\gamma _a} + {\gamma _\sigma }} \right)\varphi + {\Omega _R}\left( {2nD + D + 1} \right)/2 \label{ZE2}
\\
\dot D =  - {\gamma _D}\left( {D + 1} \right) - 4{\Omega _R}\varphi, \label{ZE3}
\end{gather}
where $n$ is the number of photons in the collective mode, $D$ is the mean population inversion of pumped atoms of the active medium, $\varphi$ describes the mean energy flow from the active medium to the collective mode, and $N_{\text{at}}$ is the number of atoms of the active medium lying in the collective mode.
We suppose that the relaxation rate of the EM field in the collective mode is equal to the characteristic relaxation rate of the modes of the plasmonic structure, i.e., $\gamma_a \simeq 10^{12} \text{s}^{-1}$.
The typical relaxation rate of the population inversion of the active medium is $\gamma_D \simeq 10^9 \text{s}^{-1}$ while the phase relaxation rate is $\gamma_{\sigma}\simeq 10^{13} \text{s}^{-1}$.
Thus, the phase relaxation rate is much larger than other relaxation rates, $\gamma_\sigma \gg \gamma_a, \gamma_D$.
For this reason, one can adiabatically eliminate $\varphi$ from Eqs.~(\ref{ZE1})--(\ref{ZE3}) and obtain the conventional rate equations
\begin{gather} 
\dot n =  - {\gamma _a}n + 2{N_{\text{at}}}\Omega _R^2\left( {2nD + D + 1} \right)/{\gamma _\sigma } \label{RE1}
\\
\dot D =  - {\gamma _D}\left( {D + 1} \right) - 4\Omega _R^2\left( {2nD + D + 1} \right)/{\gamma _\sigma } \label{RE2}
\end{gather}
As initial condition we take $D(0) = 1$ and $n(0) = 0$, i.e., all atoms are inverted and there are no photons in the mode.
The temporal dynamics of such a system is well-known \cite{SiegmanLasers}; it is shown in Fig.~\ref{fig4}.

@@

Note that the delay time, Eq.~(\ref{tdel}), resembles the dependence of the delay time of a superradiant burst \cite{gross1982superradiance, andreev1980collective,nefedkin2017badcavitySR}.
The superradiance problem and the problem that we investigate here have the same physical grounds.
Indeed, it is well known \cite{gross1982superradiance, andreev1980collective} that a system consisting of initially inverted atoms placed in a low-Q cavity, i.e., when $\gamma_a \gg \gamma_D$, exhibits a superradiant burst with the delay time, Eq.~(\ref{tdel}).

@@

It has been shown that the laser's response time can be shortened by optimizing the size of the pump beam, as a result of the consequent reduction of the delay time.
The dependence of the response time on the size of the pump beam is for following reason.
The interaction of the laser modes with the pumped active medium results in the emergence of a single collective mode.
If the size of the pump beam exceeds the propagation length of the EM waves, this collective mode is located in the pumped area.
The response time then grows logarithmically with increasing pump beam diameter.
In contrast, when the mode is outside the pumped region, the response time is approximately inversely proportional to the square of the pump beam diameter, and so increases rapidly as this diameter decreases from its critical value.
Thus, there is an optimal size of the pump beam, for which the response time is a minimum.

The existence of an optimal size gives the opportunity to reduce the energy consumption because further increase of the pump beam size does not shorten the response time.
In turn, if the area of the pump beam is small, there is the opportunity to reduce the size of the device.
Thus, the existence of an optimal pump beam size paves the way to creating ultrafast optoelectronic devices based on plasmonic lasers with low power consumption which have the potential of on-chip integration.

@@

>>>>

	The time response to an external signal is the main characteristic of optoelectronic devices which determines their maximum modulation frequency.
	The use of plasmonic structures gives the opportunity to significantly reduce the response time.
	In this paper, we study the temporal dynamics of a plasmonic distributed feedback laser consisting of a two-dimensional metallic structure coated with a layer of the active medium.
	Combining numerical simulations with an analytical evaluation we show that the response time in the large-signal modulation regime strongly depends on the size of the pump beam.
	We demonstrate that the interaction of a large number of modes of the plasmonic structure with the pumped active medium leads to the formation of a single collective mode.
	If the size of the pump beam is larger than the propagation length of electromagnetic waves in the plasmonic structure, the collective mode is localized in the pumped area.
	In this case, the response time slightly increases with the size of the pump beam.
	In the opposite case, the mode is not localized inside the pumped area and the response time grows with the narrowing of the beam.
	We demonstrate that there is a critical minimal size of the pump beam for which the modulation speed is a maximum and can achieve $1$ THz.
	The obtained result opens the way to increasing both the modulation speed and the energy efficiency of plasmonic laser devices.

@@

In our paper, we study the dynamics in the large-signal modulation regime of a plasmonic distributed feedback (DFB) laser that consists of a gold film perforated by a periodic array of holes, and is coated with a layer of an active medium.
We examine the behavior of the laser for different sizes of the pump beam.
Combining numerical simulations of the laser dynamics within the framework developed in Ref. \cite{Zyablovsky2017approach,nefedkin2018acsphot} with an analytical evaluation, we show that the response time strongly depends on the size of the pump beam.
This dependence is due to the complex structure of the modes of the DFB laser.
The interaction of the laser modes with the pumped active medium causes the emergence of a single collective mode whose properties strongly depend on the pump size.
If the pump beam is larger than the decay length of the EM field in the plasmonic structure, the collective mode is localized inside the pumped area.
The response time in this case reaches picoseconds and almost does not change with the size of the pump beam.
In the opposite case, the mode is not localized inside the pumped area and the response time grows with the narrowing of the beam.
Therefore, there is a critical minimal diameter of the pump beam, $d_{beam}^*$, required for achievement of the short response time. This diameter is about the decay length of the EM field in the plasmonic structure and for the considered plasmonic laser is $\sim 15 \mu$m. When the size of pump beam is greater than $d_{beam}^*$ the response time of the plasmonic laser is $\sim 1$ $\text{ps}$.
The obtained results confirm that plasmonic DFB lasers can have a modulation frequency exceeding several hundred GHz in the large-signal modulation regime. The optimization of the pump beam size makes it possible to increase the modulation speed as well as the energy efficiency of plasmonic DFB laser devices.

@@

\section*{Temporal dynamics of a two-dimensional plasmonic DFB laser}
A conventional plasmonic DFB laser consists of a periodic plasmonic structure, e.g., a metallic film perforated by an array of holes \cite{BeijnumPRL,DorofeenkoOptExp,MengLPR,nefedkin2018acsphot,TennerJOpt,TennerACSPhot,Zyablovsky2017Optimum,Melentiev2017nanolaser} or two-dimensional arrays of plasmonic nanoparticles \cite{ZhouNatNano,SchokkerPRB,daskalakis2018nanolett,SchokkerACSPhot,YangNatComm,YangACSNano,hakala2017lasing,wang2017band}, and an active medium.
The eigenmodes of such a system are Bloch modes which are distributed over the whole surface of the laser.
To overcome losses in the metal, an external pumping (optical or electrical) of the active medium is used.
In the case of optical pumping, the diameter of the pump beam is usually smaller than the size of the periodic plasmonic structure \cite{TennerACSPhot,ZhouNatNano,daskalakis2018nanolett}, see Fig.~\ref{fig1}a.
In such a case, the EM field is excited inside the pumped area and is formed by a large number of Bloch modes \cite{TennerACSPhot,nefedkin2018acsphot}.
To describe the multimode dynamics of the plasmonic DFB laser we use an approach which has been proposed in Ref. \cite{Zyablovsky2017approach} for considering ultrafast phenomena in dispersive dissipative media.
This approach is based on the following equations:
\begin{equation}\label{eq1}
\frac{{d{n_{jk}}}}{{dt}} =  - \left( {{\gamma _j} + {\gamma _k}} \right){n_{jk}} + i\left( {{\omega _j} - {\omega _k}} \right){n_{jk}} + \sum\limits_{m=1}^{N_{\text{atom}}} {\left( {{\Omega _{km}}{\varphi _{jm}} + \Omega _{jm}^*\varphi _{km}^*} \right)} 
\end{equation}
\begin{equation}\label{eq2}
\frac{{d{D_m}}}{{dt}} =  - {\gamma _D}\left( {1 + {D_m}} \right)  - 2\sum\limits_{j=1}^{N_{\text{mode}}} {\left( {{\Omega _{jm}}{\varphi _{jm}} + \Omega _{jm}^*\varphi _{jm}^*} \right)} 
\end{equation}
\begin{equation}\label{eq3}
\frac{{d{\varphi _{jm}}}}{{dt}} =  -\gamma_{jm}^{\varphi} {\varphi _{jm}} + i\left( {{\omega _j^a} - {\omega _\sigma }} \right){\varphi _{jm}} + \frac{{\Omega _{jm}^*}}{2}\left( {{D_m} + 1} \right) + \sum\limits_{l=1}^{N_{\text{mode}}} {\Omega _{lm}^*{n_{jl}}{D_m}} , 
\end{equation}
Here, $n_{jj}$ is the photon number in the $j$th Bloch mode, $n_{jk}$ is an interference term responsible for the flow of photons from the $k$th to the $j$th mode, when $j \neq k$.
$D_m$ is the population inversion of the $m$th atom and $\varphi _{jm} = - i \langle \hat{a} _j^{+} \hat{\sigma} _m \rangle$ describes the energy flow between the $m$th atom and $j$th mode.
The parameters $\omega _j$, $\gamma _j$ correspond to the real and imaginary parts of the Bloch modes of plasmonic structure; $\omega _{\sigma}$ and $\gamma _D$ are the atom transition frequency and the relaxation rate of the population inversion of the atoms, $\gamma _{jm}^{\varphi} = \gamma _{\sigma} + \gamma _j + \gamma _D / 2$, $\gamma _{\sigma}$ is the relaxation rate of the phase of the atom polarization.
$\Omega _{jm}$ is the Rabi constant of coupling between the $j$th mode and the $m$th atom which can be expressed as
\begin{equation}\label{eq4SM}
{\Omega _{jm}} =  - {{\bf{d}}_m} \cdot {{\bf{E}}_j}({{\bf{r}}_m})/ \hbar,
\end{equation}
where ${{\bf{d}}_m}$  is the dipole moment of the $m$th atom at the transition frequency and ${{\bf{E}}_j}({{\bf{r}}_m})$  is the electric field `per one quantum' in the $j$th mode at the position of the $m$th atom. The electric field `per one quantum' is inversely proportional to the mode volume (see Appendix A). As a result, the Rabi constant increases with the decrease of the mode volume. The Eqs.~(\ref{eq1})--(\ref{eq3}) allow for the dependence of the stimulated and spontaneous emissions rates on the Rabi constant that enables to take into account the Purcell effect of the plasmonic structure. Every atom of the active medium has own coordinates which determine the Rabi constants between the atom and the Bloch modes of the plasmonic structure, see Eq.~(\ref{eq4SM}). That makes it possible to take into account the spatial distribution of the pumping and the population inversion within the active medium.

The Eqs.~(\ref{eq1})--(\ref{eq3}) generalize the rate equations widely used in laser theory \cite{SiegmanLasers}.
In addition to the dynamics of the photon number in each mode and the population inversion of each atom, they describe energy flows between the active medium and the atoms, and the interference between the EM fields of the different modes, $n_{jk}$ \cite{Zyablovsky2017approach}.

For a quantitative description of the temporal dynamics of a plasmonic DFB laser under pulse pumping, we consider a laser based on an Au film perforated by a periodic array of nanosized holes and coated the layer of the active medium, see Fig.~\ref{fig1}. We consider that the thickness of the Au film is $100$ $\text{nm}$, whereas the diameter of the holes is $160$ $\text{nm}$ and the distance between holes is $470$ $\text{nm}$. The surface of whole periodic plasmonic structure has size of 190 $\mu$m by 1908 $\mu$m. We consider that the transition frequency of the active medium is close to the frequency of second-order Bragg grating in the plasmonic structure and the linewidth of the active medium is about 100 nm.

Similar plasmonic DFB laser have been realized in a number of experiments \cite{BeijnumPRL,TennerJOpt,TennerACSPhot} and its parameters have been well studied \cite{BeijnumPRL,TennerJOpt,TennerACSPhot,nefedkin2018acsphot}.

In the absence of the holes the eigenmodes of the plasmonic structure are TM-modes similar to surface plasmons at the metal-dielectric interface \cite{TennerJOpt,TennerACSPhot}. 
These modes have in-plane component of the magnetic field, $H_{\parallel}$, that is perpendicular to the direction of mode's propagation and both out-of-plane, $E_{\perp}$, and in-plane, $E_{\parallel}$, components of the electric field \cite{TennerJOpt,TennerACSPhot}. 
The in-plane electric field, $E_{\parallel}$, is in the direction of the mode's propagation. 
The out-of-plane component of the electric field, $E_{\perp}$, is much greater than in-plane component, $E_{\parallel}$ \cite{TennerJOpt,TennerACSPhot}.
The eigenmodes of the plasmonic structure with the holes we find within the coupled-mode theory with the experimental values of the scattering rates (see~\cite{TennerJOpt} and Section 2 in Suppl.Mat. of \cite{nefedkin2018acsphot}). The respective dispersion curves are shown \textbf{\underline{in Fig.~.}} The modes corresponding to different dispersion curves have various radiation losses and field distributions inside the unit cell of the plasmonic crystal (see Section 2 in Suppl.Mat. of \cite{nefedkin2018acsphot}). The far-field radiation from the plasmonic structure mainly arises owing to the scattering of $H_{\parallel}$ by the holes~\cite{TennerJOpt}. 
For this reason, the modes that have a minimum value of $H_{\parallel}$ at the holes have smaller radiation losses than the modes that have a maximum value of $H_{\parallel}$ at the holes.
In the continuous-wave regime the lasing takes place at the modes with small radiation losses and high quality factor lied in 2nd dispersion curve, \textbf{\underline{see Fig.~}} \cite{TennerJOpt,nefedkin2018acsphot}.

In numerical simulation we allow for several hundreds the Bloch modes of the plasmonic structure lied in different dispersion curves. The step of quantization of the Bloch wavevectors is determined by the size of the periodic plasmonic structure \cite{nefedkin2018acsphot} and Suppl.Mat. of \cite{nefedkin2018acsphot}.

We consider the response of the system to ultrashort pulse pumping of the active medium.
Assuming that the time of the external pump pulse lies in the femtosecond range and is much shorter than all characteristic times of the plasmonic DFB laser, in our model the ultrashort pulse pumping is described as the initial condition for the $D_m$ in Eqs.~(\ref{eq1})--(\ref{eq3}). We consider that in an initial moment the population inversions of the atoms whose coordinate lie inside the pump beam are expressed as
\begin{equation}\label{eq5SM}
{\left(D_m+1\right)/2} = {P_0 \Delta t / \left(n_c S h \hbar \omega \right)}. 
\end{equation}
Here $\left(D_m+1\right)/2$ is a population of the upper level of the atom; $P_0$ is a pump power; $\Delta t$ is a duration of the pump pulse; $n_c$ is a concentration of the active atoms; $S$ is an area of the pump beam; $h$ is the thickness of Au film; $\omega$ is a frequency of the pump pulse. Also, we consider that the population inversions of the atoms whose coordinate lie outside the pump beam are $-1$ in the initial moment.

@@

\begin{figure}[h]
	\centering
	\begin{subfigure}[h]{0.32\linewidth}
		\includegraphics[width=\linewidth]{Fig3a.png}
%		\caption{}
	\end{subfigure}
	\begin{subfigure}[h]{0.32\linewidth}
		\includegraphics[width=\linewidth]{Fig3b.png}
%		\caption{}
	\end{subfigure}
	\begin{subfigure}[h]{0.32\linewidth}
		\includegraphics[width=\linewidth]{Fig3c.png}
%		\caption{}
	\end{subfigure}
	\\
	\begin{subfigure}[h]{0.32\linewidth}
		\includegraphics[width=\linewidth]{Fig3d.png}
%		\caption{}
	\end{subfigure}
	\begin{subfigure}[h]{0.32\linewidth}
		\includegraphics[width=\linewidth]{Fig3e.png}
%		\caption{}
	\end{subfigure}
	\begin{subfigure}[h]{0.32\linewidth}
		\includegraphics[width=\linewidth]{Fig3f.png}
%		\caption{}
	\end{subfigure}
	\caption{Distribution of the EM field intensity on the surface of plasmonic DFB laser when diameter of the pump beam $d_{\text{beam}}<d_{\text{beam}}^*$ (a)-(c) in time moment 0.5 ps (a); 1.75 ps (b); 2.5 ps (c). Distribution of the EM field intensity on the surface of plasmonic DFB laser when diameter of the pump beam $d_{\text{beam}}>d_{\text{beam}}^*$ (d)-(f) in time moment 0.32 ps (d); 0.541 ps (e); 1.5 ps (f). These time moments correspond to circles on the red curve and the black curve in Fig.\ref{fig2}a, respectively. The dashed circles in Figs. 3 denote the boundaries of the pump beam.}
	\label{fig3}	
\end{figure}

@@

The existence of an optimal size of the pump beam can be explained with the example of a single-mode model. To justify the single-mode approximation, we calculate the distribution of the electric field intensity over the area of the plasmonic DFB laser at different moments of time, see Fig.~\ref{fig3}.
Although, in our simulation we take into account the several hundreds of the Bloch modes of the plasmonic structure which jointly form the EM field distribution in the active medium we can see that the EM field distribution practically does not change in the first and second stage of the system response. Another words, there is dynamical balance between the relaxation processed in the plasmonic structure, the propagation of the EM field and the stimulated amplification of the EM field in the active medium. When the diameter of the pump beam is larger than the critical value, $d_{\text{beam}} > d_{\text{beam}}^*$, the electric field is located inside the pumped region of the active medium, see Figs.~\ref{fig3}(d)--(f).
Whereas when $d_{\text{beam}} < d_{\text{beam}}^*$, the main part of the electric field is outside the pumped area, see Figs.~\ref{fig3}(a)--(c).
However, in both cases, the spatial distribution of the electric field intensity remains almost unchanged during its temporal evolution.

We use the fact of the constancy of the EM field distribution for construction of the single-mode model. For this purpose, we consider that there is single collective mode whose the EM field distribution coincide with the EM field distribution found in the numerical simulation multi-mode problem. Furthermore, we assume that the pumped active medium interacts only with such a single collective mode. Within of these approximations, the system dynamics can be approximately described by the following system of equations \cite{SiegmanLasers}:
\begin{gather} 
\dot n =  - 2{\gamma _a}n + 2{N_{\text{at}}}{\Omega _R}\varphi \label{ZE1}
\\
\dot \varphi = - \left( {{\gamma _a} + {\gamma _\sigma }} \right)\varphi + {\Omega _R}\left( {2nD + D + 1} \right)/2 \label{ZE2}
\\
\dot D =  - {\gamma _D}\left( {D + 1} \right) - 4{\Omega _R}\varphi, \label{ZE3}
\end{gather}
where $n$ is the number of photons in the collective mode, $D$ is the mean population inversion of pumped atoms of the active medium, $\varphi$ describes the mean energy flow from the active medium to the collective mode, and $N_{\text{at}}$ is the number of atoms of the active medium lying in the collective mode.
We suppose that the relaxation rate of the EM field in the collective mode is equal to the characteristic relaxation rate of the modes of the plasmonic structure, i.e., $\gamma_a \simeq 10^{12} \text{s}^{-1}$.
The typical relaxation rate of the population inversion of the active medium is $\gamma_D \simeq 10^9 \text{s}^{-1}$ while the phase relaxation rate is $\gamma_{\sigma}\simeq 10^{13} \text{s}^{-1}$.
Thus, the phase relaxation rate is much larger than other relaxation rates, $\gamma_\sigma \gg \gamma_a, \gamma_D$.
For this reason, one can adiabatically eliminate $\varphi$ from Eqs.~(\ref{ZE1})--(\ref{ZE3}) and obtain the conventional rate equations
\begin{gather} 
\dot n =  - {\gamma _a}n + 2{N_{\text{at}}}\Omega _R^2\left( {2nD + D + 1} \right)/{\gamma _\sigma } \label{RE1}
\\
\dot D =  - {\gamma _D}\left( {D + 1} \right) - 4\Omega _R^2\left( {2nD + D + 1} \right)/{\gamma _\sigma } \label{RE2}
\end{gather}
As initial condition we take $D(0) = 1$ and $n(0) = 0$, i.e., all atoms are inverted and there are no photons in the mode.
The temporal dynamics of such a system is well-known \cite{SiegmanLasers}; it is shown in Fig.~\ref{fig4}.

@@

@@

It has been shown that the laser's response time can be shortened by optimizing the size of the pump beam, as a result of the consequent reduction of the delay time.
The dependence of the response time on the size of the pump beam is for following reason.
The interaction of the laser modes with the pumped active medium results in the emergence of a single collective mode.
If the size of the pump beam exceeds the propagation length of the EM waves, this collective mode is located in the pumped area.
The response time then grows logarithmically with increasing pump beam diameter.
In contrast, when the mode is outside the pumped region, the response time is approximately inversely proportional to the square of the pump beam diameter, and so increases rapidly as this diameter decreases from its critical value.

Thus, there is a critical minimal diameter of the pump beam, $d_{beam}^*$, required for achievement of the short response time. This diameter is about the decay length of the EM field in the plasmonic structure and for the considered plasmonic laser is $\sim 15 \mu$m. The existence of the critical pump beam size gives the opportunity to reduce the energy consumption because further increase of the pump beam size does not shorten the response time.
In turn, if the area of the pump beam is small, there is the opportunity to reduce the size of the device.
Thus, the obtained results pave the way to creating ultrafast optoelectronic devices based on plasmonic lasers with low power consumption which have the potential of on-chip integration.

It is interesting that the delay time, Eq.~(\ref{tdel}), resembles the dependence of the delay time of a superradiant burst \cite{gross1982superradiance, andreev1980collective,nefedkin2017badcavitySR}.
Indeed, it is well known \cite{gross1982superradiance, andreev1980collective} that a system consisting of initially inverted atoms placed in a low-Q cavity, i.e., when $\gamma_a \gg \gamma_D$, exhibits a superradiant burst with the delay time, Eq.~(\ref{tdel}). In the plasmonic cavity the inequality $\gamma_a \gg \gamma_D$ always is hold true and, for this reason, the superradiance problem and the problem that we investigate here have the same physical grounds. We believe that further study of connection between these problems is interesting.

@@

\section*{Appendix A}
To normalize the electric field ${{\bf{E}}_j}({{\bf{r}}_m})$, we equate the energy of one quantum to the energy of the electric field in the resonator:
\begin{equation}
\hbar {\omega _j} = \frac{1}{{8\pi }}\int\limits_V {\left( {{{\left. {\frac{{\partial {\mathop{\rm Re}\nolimits} \left( {\varepsilon(\omega, \textbf{r}) \omega } \right)}}{{\partial \omega }}} \right|}_{\omega  = {\omega _j}}}{{\left| {{{\bf{E}}_j}\left( {\bf{r}} \right)} \right|}^2} + {{\left| {{{\bf{H}}_j}\left( {\bf{r}} \right)} \right|}^2}} \right){d^3}{\bf{r}}} \tag{A.1}
\end{equation}
where $\varepsilon (\omega, \textbf{r})$  is the dielectric constant of the active plasmonic structure.

In a two-dimensional periodic plasmonic structure whose surface is parallel to the $xy$-plane, the electric and magnetic fields satisfy the Bloch condition,
\begin{equation}
{{\bf{E}}_j}\left( {\bf{r}} \right) = {A_{0j}}\exp \left( {i{k_{xj}}x + i{k_{yj}}y} \right){{\bf{e}}_j}\left( {x,y,z} \right) \tag{A.2}
\end{equation}
\begin{equation}
{{\bf{H}}_j}\left( {\bf{r}} \right) = {A_{0j}}\exp \left( {i{k_{xj}}x + i{k_{yj}}y} \right){{\bf{h}}_j}\left( {x,y,z} \right), \tag{A.3}
\end{equation}
where $A_{0j}$ is a real normalization constant for \textit{j}th mode; ${{\bf{e}}_j}\left( {x,y,z} \right) = {{\bf{e}}_j}\left( {x + {L_x},y + {L_y},z} \right)$  and ${{\bf{h}}_j}\left( {x,y,z} \right) = {{\bf{h}}_j}\left( {x + {L_x},y + {L_y},z} \right)$, where $L_x$ and $L_y$ are periods of the plasmonic structure along $x$-axis and $y$-axis, respectively.

Suppose that the area under investigation contains an integral number $N$ of plasmonic lattice cells.
Then Eq. (A.1) may be rewritten as
\begin{equation}
\hbar {\omega _j} =\frac{N}{{8\pi }} A_{0j}^2 \int\limits_{{V_{cell}}} {\left( {{{\left. {\frac{{\partial {\mathop{\rm Re}\nolimits} \left( {\varepsilon(\omega, \textbf{r}) \omega } \right)}}{{\partial \omega }}} \right|}_{\omega  = {\omega _i}}}{{\left| {{{\bf{e}}_j}\left( {\bf{r}} \right)} \right|}^2} + {{\left| {{{\bf{h}}_j}\left( {\bf{r}} \right)} \right|}^2}} \right){d^3}{\bf{r}}} \tag{A.4}
\end{equation}
where ${V_{cell}}$  is the volume of the cell of the plasmonic lattice. As a result, we obtain that the normalization constant $A_{0j}$ is expressed as
\begin{equation}
A_{0j} =\sqrt{\frac{\hbar {\omega _j}}{\frac{N}{{8\pi }} \int\limits_{{V_{cell}}} {\left( {{{\left. {\frac{{\partial {\mathop{\rm Re}\nolimits} \left( {\varepsilon(\omega, \textbf{r}) \omega } \right)}}{{\partial \omega }}} \right|}_{\omega  = {\omega _i}}}{{\left| {{{\bf{e}}_j}\left( {\bf{r}} \right)} \right|}^2} + {{\left| {{{\bf{h}}_j}\left( {\bf{r}} \right)} \right|}^2}} \right){d^3}{\bf{r}}}}} \tag{A.5}
\end{equation}
